---
title: "Functional label propagation"
author: "James C. Kosmopoulos"
date: "`r Sys.Date()`"
output:
  github_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 6, dpi = 150)

library("arrow");packageVersion("arrow")
library("tidyverse");packageVersion("tidyverse")
library("cowplot");packageVersion("cowplot")
library("scales");packageVersion("scales")
library("patchwork");packageVersion("patchwork")
library("jsonlite");packageVersion("jsonlite")

PROP_DIR  <- file.path("./tables/propagation")
NOVEL_DIR <- file.path("./tables/novel_avgs")
PLOT_DIR  <- file.path("./plots/functional_propagation")
dir.create(PLOT_DIR, showWarnings = FALSE, recursive = TRUE)

SCHEMA_VERSION <- "propagation-1.2"
MANIFEST <- jsonlite::fromJSON(file.path(PROP_DIR, "manifest.json"))
stopifnot(identical(MANIFEST$schema_version, SCHEMA_VERSION))
EVAL <- jsonlite::fromJSON(file.path(PROP_DIR, "evaluation_summary.json"))
TARGET <- MANIFEST$primary_target

prop  <- function(name) read_parquet(file.path(PROP_DIR, paste0(name, ".parquet")))
novel <- function(name) read_parquet(file.path(NOVEL_DIR, paste0(name, ".parquet")))
save_plot_both <- function(p, stem, w, h, dpi = 600, bg = "white") {
  ggsave(file.path(PLOT_DIR, paste0(stem, ".png")), p, width = w, height = h, units = "in", dpi = dpi, bg = bg)
  invisible(p)
}
```

# Palette and theme

```{r palettes}
# level colors run dark to light with depth of information, so specific reads as the strongest claim
level_lvl <- c("specific", "L1", "category", "unassigned")
level_lab <- c("specific" = "Specific function", "L1" = "L1 function category",
               "category" = "Broad category", "unassigned" = "No label at target")
level_pal <- c("specific" = "#1a9850", "L1" = "#74add1", "category" = "#fdae61", "unassigned" = "grey70")

category_pal <- c("metabolic" = "#2c7fb8", "physiological" = "#7fbc41", "regulatory" = "#d95f0e")
rank_pal <- c("calibrated" = "#1a9850", "d1_only" = "grey55")
rank_lab <- c("calibrated" = "Calibrated probability", "d1_only" = "Embedding distance only")

# tier scales: cluster tier purple, embedding tier the same green the calibrated curves use
reach_lvl <- c("both", "cluster_only", "embedding_only", "neither")
reach_lab <- c("both" = "Both tiers", "cluster_only" = "Cluster tier only",
               "embedding_only" = "Embedding tier only", "neither" = "Neither tier")
method_lvl <- c("cluster_only", "embedding_only", "tiered")
method_lab <- c("cluster_only" = "Cluster tier alone", "embedding_only" = "Calibrated embedding alone",
                "tiered" = "Deployed tiered rule")
method_pal <- c("cluster_only" = "#762a83", "embedding_only" = "#1a9850", "tiered" = "#0f3b5f")
source_lvl <- c("cluster", "embedding", "none")
source_lab <- c("cluster" = "Cluster tier", "embedding" = "Calibrated embedding tier",
                "none" = "No label at target")
source_pal <- c("cluster" = "#762a83", "embedding" = "#1a9850", "none" = "grey70")

base_theme <- theme_cowplot(font_size = 12) +
  theme(axis.text = element_text(size = 11), axis.title = element_text(size = 13),
        legend.title = element_text(size = 13, face = "bold"), legend.text = element_text(size = 11),
        strip.text = element_text(size = 12, face = "bold"),
        legend.key.size = unit(0.42, "cm"), panel.grid.major = element_blank())
theme_set(base_theme)
```

# Method comparison at the deployed target

```{r tier-headline, include=FALSE}
TIER_CMP <- prop("fig_tier_coverage_precision") %>%
  filter(abs(target - TARGET) < 1e-9, level == "any") %>%
  select(method, n_report, coverage, precision, recall)
stopifnot(setequal(TIER_CMP$method, c("cluster_only", "embedding_only", "tiered")))
tcmp <- function(m, col, acc = 0.01) {
  v <- TIER_CMP[[col]][TIER_CMP$method == m]
  stopifnot(length(v) == 1)
  percent(v, accuracy = acc)
}
PREC_GIVEUP <- TIER_CMP$precision[TIER_CMP$method == "cluster_only"] -
  TIER_CMP$precision[TIER_CMP$method == "tiered"]
SWEEP_AT <- prop("fig_target_sweep_heldout") %>% filter(abs(target - TARGET) < 1e-9)
swp <- function(lv, acc = 0.01) {
  v <- SWEEP_AT$realized_precision[SWEEP_AT$level == lv]
  stopifnot(length(v) == 1)
  percent(v, accuracy = acc)
}
TIER_OVL <- prop("fig_tier_overlap")
ovlv <- function(lvl, grp, col) {
  v <- TIER_OVL[[col]][TIER_OVL$level == lvl & TIER_OVL$reachable_by == grp]
  stopifnot(length(v) == 1)
  v
}
TIER_STRAT <- prop("fig_tier_by_identity")
tstv <- function(st, m, col) {
  v <- TIER_STRAT[[col]][TIER_STRAT$stratum == st & TIER_STRAT$method == m]
  stopifnot(length(v) == 1)
  v
}
TIER_APP <- prop("fig_tier_assignment_breakdown") %>%
  filter(scope == "sequence_similarity_invisible", abs(target - TARGET) < 1e-9)
N_HI <- sum(TIER_APP$n)
appv <- function(src) sum(TIER_APP$n[TIER_APP$label_source == src])
```

# Propagation coverage and precision at different sequence identity thresholds, for each propagation tier

```{r plot-coverage-precision-by-seq-id, fig.width=16, fig.height=4, fig.show="hide"}
tstrat <- prop("fig_tier_by_identity") %>%
  mutate(stratum = factor(stratum, levels = c("<50%", "50-90%", ">=90%", "all"),
                          labels = c("< 50%", "50-90%", ">= 90%", "All")),
         method = factor(method, levels = method_lvl))
xlabs <- tstrat %>% distinct(stratum, n_report) %>%
  mutate(lab = paste0(stratum, "\nn = ", comma(n_report))) %>% arrange(stratum)
dg <- position_dodge(width = 0.78)

p_prop_cov_precision <- ggplot(tstrat, aes(stratum, coverage, fill = method)) +
  geom_col(position = dg, width = 0.72, alpha = 0.45, color = NA) +
  geom_line(aes(y = precision, color = method, group = method), position = dg, linewidth = 0.7) +
  geom_point(aes(y = precision, color = method), position = dg, size = 2.4) +
  scale_fill_manual(values = method_pal, labels = method_lab, name = NULL) +
  scale_color_manual(values = method_pal, labels = method_lab, name = NULL) +
  scale_x_discrete(labels = xlabs$lab) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1),
                     expand = expansion(c(0, 0.02))) +
  labs(x = "Seq. identity to nearest training protein",
       y = "Coverage (bars) and\nprecision (points)") +
  theme(legend.position = "bottom", legend.direction = "vertical", legend.location = "plot",
        plot.subtitle = element_text(size = 9.5), axis.text.x = element_text(size = 9.5),
        axis.title.y = element_text(size = 11.5),
        plot.margin = margin(t = 12, r = 4, b = 2, l = 4, unit = "pt"))
print(p_prop_cov_precision)
```

# Which propagation tier could reach each held-out protein

```{r plot-propagation-reach, fig.width=16, fig.height=4, fig.show="hide"}
ovl <- prop("fig_tier_overlap")
stopifnot(nrow(ovl) == 12, setequal(ovl$reachable_by, reach_lvl))
N_REPORT <- round(unique(ovl$n / (ovl$pct_of_slice / 100))[1])
pc1 <- function(x) sprintf("%.1f", 100 * x)

p_propagation_reach <- ovl %>%
  mutate(level = factor(level, levels = c("specific", "L1", "category")),
         reachable_by = factor(reachable_by, levels = rev(reach_lvl)),
         lab = paste0(comma(n), " (", pc1(pct_of_slice / 100), "%)\n",
                      case_when(
                        reachable_by == "both" ~
                          paste0("C ", pc1(tier1_precision), " | E ", pc1(tier2_precision)),
                        reachable_by == "cluster_only" ~
                          paste0("C ", pc1(tier1_precision), " | 1NN ", pc1(tier2_precision)),
                        reachable_by == "embedding_only" ~ paste0("E ", pc1(tier2_precision)),
                        TRUE ~ paste0("1NN ", pc1(tier2_precision))))) %>%
  ggplot(aes(level, reachable_by, fill = pct_of_slice)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = lab), size = 3.1, lineheight = 0.95) +
  scale_fill_gradient(low = "#f7f7f7", high = "#9ecae1", name = "Share of held-out proteins",
                      labels = function(x) paste0(x, "%"), guide = guide_colorbar(barwidth = 8)) +
  scale_x_discrete(labels = c("specific" = "Specific", "L1" = "L1 category",
                              "category" = "Broad category")) +
  scale_y_discrete(labels = reach_lab) +
  labs(x = paste0("Level at which a tier could assign, ", comma(N_REPORT), " held-out proteins"),
       y = NULL,
       subtitle = paste0("C = cluster tier precision; E = admitted embedding precision;\n1NN = precision the uncalibrated neighbor would have delivered")) +
  theme(
    legend.position = "bottom",
    plot.subtitle = element_text(size = 9.5),
    axis.text.y = element_text(size = 10.5)
  )
print(p_propagation_reach)
```

# Validation controls of label propagation

```{r plot-propagation-validation-controls}
ctrl <- prop("fig_validation_controls")
ctrl$panel <- gsub("Contribution of finetuning", "Embedding space", ctrl$panel)
ctrl <- ctrl %>%
  mutate(panel = factor(panel, levels = c("Unseen held-out proteins", "Unseen genomic context",
                                          "Specificity", "Embedding space"),
                        labels = c("Unseen held-out proteins\n(category precision)",
                                   "Unseen genomic context\n(category precision)",
                                   "Specificity\n(confident-call rate)",
                                   "Embedding space\n(category precision)")),
         condition = fct_rev(fct_inorder(condition)))

p_ctrl <- ggplot(ctrl, aes(value, condition, fill = panel)) +
  geom_col(width = 0.68, color = "black", linewidth = 0.25) +
  geom_text(aes(label = ifelse(value < 0.01, sprintf("%.4f%%", 100 * value),
                               sprintf("%.1f%%", 100 * value))),
            hjust = -0.08, size = 3) +
  facet_grid(panel ~ ., scales = "free_y", space = "free_y", switch = "y") +
  scale_x_continuous(labels = percent, expand = expansion(mult = c(0, 0.12)), breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  scale_fill_brewer(palette = "Set2", guide = "none") +
  labs(x = "Proportion of proteins", y = NULL) +
  theme(strip.placement = "outside", strip.text.y.left = element_text(angle = 0, size = 9, face = "bold"),
        axis.text.y = element_text(size = 9), panel.spacing.y = unit(0.25, "lines"))

print(p_ctrl)
```

# What propagation delivers on the applied auxiliary viral genes across prediction targets

```{r plot-propagation-composition-by-precision-target, fig.width=6, fig.height=6, fig.show="hide"}
hi_sweep <- prop("fig_target_sweep_applied")

p_prop_comp_by_target <- hi_sweep %>%
  mutate(level = factor(level, levels = level_lvl)) %>%
  ggplot(aes(target, pct, fill = level)) +
  geom_col(width = 0.045) +
  scale_fill_manual(values = level_pal, labels = level_lab, name = NULL) +
  scale_y_continuous(labels = function(x) paste0(x, "%"), expand = expansion(c(0, 0.02))) +
  scale_x_continuous(breaks = unique(hi_sweep$target), labels = percent_format(accuracy = 1)) +
  labs(x = "Precision target", y = "Sequence-similarity invisible AVGs") +
  theme(
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    legend.position = "bottom",
    legend.justification = "center",
    legend.location = "plot",
    legend.direction = "vertical"
    ) +
  guides(fill = guide_legend(nrow = 3))

print(p_prop_comp_by_target)
```

# Assignment depth of each propagation tier for sequence similarity invisible AVG slices

```{r plot-tier-assignment-depth, fig.width=8, fig.height=3.5, fig.show="hide"}
SUB_LVL <- c("All invisible AVGs", "Already identified", "Unresolved")
nm <- novel("novel_master_per_protein") %>% select(known_function, label_source, final_level)
app <- bind_rows(
    nm %>% mutate(subset = SUB_LVL[1]),
    nm %>% filter(known_function)  %>% mutate(subset = SUB_LVL[2]),
    nm %>% filter(!known_function) %>% mutate(subset = SUB_LVL[3])) %>%
  count(subset, final_level, label_source, name = "n") %>%
  group_by(subset) %>% mutate(n_subset = sum(n), frac = n / n_subset) %>% ungroup() %>%
  mutate(final_level = factor(final_level, levels = level_lvl),
         label_source = factor(label_source, levels = source_lvl))

stopifnot(identical(
  app %>% filter(subset == SUB_LVL[1]) %>%
    transmute(level = as.character(final_level), label_source = as.character(label_source), n) %>%
    arrange(level, label_source) %>% pull(n),
  TIER_APP %>% transmute(level, label_source, n = as.integer(n)) %>%
    arrange(level, label_source) %>% pull(n)))

sub_labs <- app %>% distinct(subset, n_subset) %>%
  mutate(lab = paste0(subset, "\nn = ", comma(n_subset)))
sub_labs <- setNames(sub_labs$lab, sub_labs$subset)
app <- app %>% mutate(subset = factor(subset, levels = SUB_LVL))

p_tier_assignment_depth <- ggplot(app, aes(frac, fct_rev(final_level), fill = label_source)) +
  geom_col(width = 0.72, color = "black", linewidth = 0.2) +
  geom_text(aes(label = ifelse(frac >= 0.02, comma(n), ""),
                ),
            position = position_stack(vjust = 0.5), size = 2.6) +
  facet_wrap(~ subset, nrow = 1, labeller = as_labeller(sub_labs)) +
  scale_fill_manual(values = source_pal, labels = source_lab, name = NULL) +
  scale_y_discrete(labels = c("specific" = "Specific", "L1" = "L1", "category" = "Broad",
                              "unassigned" = "None")) +
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(c(0.03, 0.04)),
                     breaks = c(0, 0.2, 0.4)) +
  coord_cartesian(clip = "off") +
  coord_flip() +
  labs(y = "Assignment depth at the deployed target", x = "Share of the subset") +
  theme(legend.position = "bottom", legend.location = "plot", legend.direction = "vertical",
        strip.text = element_text(size = 8.8), strip.clip = "off",
        plot.subtitle = element_text(size = 9),
        panel.spacing.x = unit(0.7, "lines"), axis.text.x = element_text(size = 9))

print(p_tier_assignment_depth)
```

# Precision of the uncalibrated neighbor against sequence identity

```{r plot-precision-vs-seq-identity, fig.width=5, fig.height=6, fig.show="hide"}
joint <- prop("fig_identity_by_distance")
strat <- prop("fig_ladder_by_identity")

p_precision_vs_seq_id <- strat %>%
  pivot_longer(c(category, L1, specific), names_to = "level", values_to = "precision") %>%
  mutate(level = factor(level, levels = c("specific", "L1", "category"))) %>%
  ggplot(aes(band, precision, color = level, group = level)) +
  geom_line(linewidth = 0.9) + geom_point(size = 2) +
  scale_color_manual(values = level_pal, labels = level_lab, name = NULL) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(x = "Seq. identity to nearest training protein", y = "Raw embedding 1-NN precision") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    legend.direction = "vertical",
    legend.text = element_text(size = 9.5),
    legend.key.size = unit(0.32, "cm"),
    plot.margin = margin(t = 2, r = 4, b = 8, l = 4, unit = "pt")
  )

print(p_precision_vs_seq_id)
```

# Realized precision vs. precision target at each propagation level

```{r propagation-levels-realized-vs-target-precision, fig.width=5, fig.height=6, fig.show="hide"}
sweep <- prop("fig_target_sweep_heldout")

p_realized_v_target <- sweep %>%
  filter(!is.na(realized_precision), level != "unassigned") %>%
  mutate(level = factor(level, levels = level_lvl)) %>%
  ggplot(aes(target, realized_precision, color = level)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, color = "grey60") +
  geom_line(linewidth = 0.8) + geom_point(size = 2) +
  scale_color_manual(values = level_pal, labels = level_lab, name = NULL) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "Precision target", y = "Realized precision") +
  theme(legend.position = "bottom", legend.direction = "vertical", legend.location = "plot", plot.margin = margin(l = 0, r =))

print(p_realized_v_target)
```

# Precision vs. propagation coverage at each level

```{r plot-precision-vs-coverage, fig.width=12, fig.height=6, fig.show="hide"}
rc <- prop("fig_risk_coverage")

p_precison_vs_coverage <- rc %>%
  mutate(level = factor(level, levels = c("specific", "L1", "category")),
         ranking = factor(ranking, levels = names(rank_pal))) %>%
  ggplot(aes(coverage, precision, color = ranking)) +
  geom_hline(yintercept = TARGET, linetype = 2, color = "grey60") +
  geom_line(linewidth = 0.9) +
  facet_wrap(~ level, nrow = 1, labeller = as_labeller(
    c("specific" = "Specific function", "L1" = "L1 function category", "category" = "Broad category"))) +
  scale_color_manual(values = rank_pal, labels = rank_lab, name = NULL) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(x = "Coverage", y = "Precision") +
  theme(legend.position = "bottom")

print(p_precison_vs_coverage)
```

# Accuraacy of the calibrated probabilities at each level

```{r plot-accuracy-vs-probability, fig.width=12, fig.height=6, fig.show="hide"}
rel <- prop("fig_reliability")
ece <- tibble(level = names(EVAL$calibration_ece), ece = unlist(EVAL$calibration_ece)) %>%
  mutate(lab = paste0("ECE = ", format(round(ece, 4), nsmall = 4)))

p_accuracy_vs_prob <- rel %>%
  mutate(level = factor(level, levels = c("specific", "L1", "category"))) %>%
  ggplot(aes(confidence, accuracy)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, color = "grey60") +
  geom_point(aes(size = n), color = "#1a9850", alpha = 0.8) +
  geom_line(color = "#1a9850", linewidth = 0.6) +
  geom_text(data = ece %>% mutate(level = factor(level, levels = c("specific", "L1", "category"))),
            aes(x = 0.03, y = 0.96, label = lab), hjust = 0, size = 4, inherit.aes = FALSE) +
  facet_wrap(~ level, nrow = 1, labeller = as_labeller(
    c("specific" = "Specific function", "L1" = "L1 function category", "category" = "Broad category"))) +
  scale_size_continuous(range = c(0.8, 4.5), guide = "none") +
  scale_x_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(x = "Calibrated probability", y = "Observed accuracy")

print(p_accuracy_vs_prob)
```

# Composite figure

```{r composite, fig.width=16, fig.height=16}
p_row1 <- ((p_prop_cov_precision | p_propagation_reach | p_ctrl) + plot_layout(widths = c(4.25, 6, 5.75)))
p_row2 <- ((p_prop_comp_by_target | p_tier_assignment_depth | p_precision_vs_seq_id) + plot_layout(widths = c(4.25, 7.25, 4)))
p_row3 <- ((p_realized_v_target | p_precison_vs_coverage | p_accuracy_vs_prob) + plot_layout(widths = c(2, 7, 7))) &
  theme(axis.text.x = element_text(size = 10))

p_comp <- (p_row1 / p_row2 / p_row3) + plot_annotation(tag_levels = "A")
p_comp <- p_comp +
  plot_layout(heights = c(1, 1, 1)) &
  theme(plot.tag = element_text(size = 16, face = "bold", hjust = 0, vjust = 1),
        plot.tag.position = "topleft", plot.tag.location = "margin")

save_plot_both(p_comp, "functional_propagation_supplement", w = 16, h = 16)
print(p_comp)
```

# Supplemental table

```{r supp-table}
supp <- prop("fig_method_summary")
knitr::kable(supp, digits = 4)
```

# Source Data and Supplemental Tables

```{r source-data-and-supp-tables}
SRC_DIR <- file.path(PROP_DIR, "source_data")
dir.create(SRC_DIR, showWarnings = FALSE, recursive = TRUE)

readr::write_csv(tstrat,   file.path(SRC_DIR, "edfig9A_tier_by_identity.csv"))
readr::write_csv(ovl,      file.path(SRC_DIR, "edfig9B_tier_reach_overlap.csv"))
readr::write_csv(ctrl,     file.path(SRC_DIR, "edfig9C_validation_controls.csv"))
readr::write_csv(hi_sweep, file.path(SRC_DIR, "edfig9D_applied_target_sweep.csv"))
readr::write_csv(app,      file.path(SRC_DIR, "edfig9E_applied_label_source.csv"))
readr::write_csv(strat,    file.path(SRC_DIR, "edfig9F_precision_by_identity_band.csv"))
readr::write_csv(sweep,    file.path(SRC_DIR, "edfig9G_target_sweep.csv"))
readr::write_csv(rc,       file.path(SRC_DIR, "edfig9H_risk_coverage.csv"))
readr::write_csv(rel,      file.path(SRC_DIR, "edfig9I_reliability.csv"))

readr::write_csv(joint,    file.path(SRC_DIR, "edfig9_extra_identity_by_distance.csv"))
cat(sprintf("source data written to %s\n", SRC_DIR))
print(tibble::tibble(
  file = list.files(SRC_DIR, pattern = "^edfig9"),
  rows = vapply(list.files(SRC_DIR, pattern = "^edfig9", full.names = TRUE),
                function(p) nrow(readr::read_csv(p, show_col_types = FALSE)), integer(1))))

supp_dir <- PROP_DIR

lad  <- prop("fig_ladder_micro_macro"); ladf <- prop("fig_ladder_by_frequency")
ladi <- prop("fig_ladder_by_identity"); rcs <- prop("fig_risk_coverage_summary")
ops  <- prop("fig_operating_points")

summary_table <- bind_rows(
  lad %>% pivot_longer(c(micro, macro), names_to = "metric", values_to = "value") %>%
    transmute(section = "performance", level, stratum_type = "overall", stratum = "all",
              metric = paste0(metric, "_precision"), value, n),
  lad %>% transmute(section = "performance", level, stratum_type = "overall", stratum = "all",
                    metric = "n_distinct_labels", value = as.numeric(n_labels), n),
  ladf %>% transmute(section = "performance", level, stratum_type = "reference_frequency_tercile",
                     stratum = freq_tercile, metric = "precision", value = precision, n),
  ladi %>% pivot_longer(c(category, L1, specific), names_to = "level", values_to = "value") %>%
    transmute(section = "performance", level, stratum_type = "identity_to_training",
              stratum = band, metric = "precision", value, n),
  tibble(level = names(EVAL$calibration_ece), value = unlist(EVAL$calibration_ece)) %>%
    transmute(section = "performance", level, stratum_type = "overall", stratum = "all",
              metric = "expected_calibration_error", value, n = NA_integer_),
  ops %>% filter(abs(target - MANIFEST$primary_target) < 1e-9) %>%
    pivot_longer(c(min_seq_id, p_threshold, coverage, precision),
                 names_to = "metric", values_to = "value") %>%
    transmute(section = "performance", level, stratum_type = "operating_point",
              stratum = sprintf("target_%.2f_%s", target, tier), metric, value, n = n_report),
  rcs %>% pivot_longer(c(aurc, `cov@0.90`, `cov@0.80`, `cov@0.70`),
                       names_to = "metric", values_to = "value") %>%
    transmute(section = "performance", level, stratum_type = "risk_coverage",
              stratum = ranking, metric, value, n = NA_integer_),
  prop("fig_validation_controls") %>%
    transmute(section = "validation_control", level = NA_character_, stratum_type = panel,
              stratum = condition, metric, value, n = NA_integer_),
  prop("fig_target_sweep_heldout") %>%
    pivot_longer(c(rate_heldout, realized_precision), names_to = "metric", values_to = "value") %>%
    transmute(section = "target_sweep", level, stratum_type = "held-out test proteins",
              stratum = sprintf("target_%.2f", target), metric, value, n),
  prop("fig_target_sweep_applied") %>%
    transmute(section = "target_sweep", level, stratum_type = "sequence-similarity invisible AVGs",
              stratum = sprintf("target_%.2f", target), metric = "rate", value = pct / 100, n)
) %>% relocate(section, level, stratum_type, stratum, metric, value, n)

readr::write_csv(summary_table, file.path(PROP_DIR, "propagation_evaluation.csv"))
```

# Session info

```{r session-info}
sessionInfo()
```
